Molecular Gas Dynamics of the lensed wet-merger RXSJ1131-1231 at $z$=0.654

### 2016-06-28 WRITEUP/ms.tex
finished discussion

Alternatively, the relatively low SFE maybe intrinsic to RXJ1131.
In that case,  RXJ1131 may evolve into a disc galaxy with a small bulge component
in the context of theoretical simulations by \citet{Hopkins09a}, who suggest 
that the disc component in a progenitor galaxy can survive post-merging
by lowering the \SF efficiency, enabling a higher gas fraction to be retained and 
distributed over a larger extent.
% 1:1 gas-riich mergers can yield disc-dominated remnants, and 1:3-1:4 ratio major mergers can yield B/T < 0.2.

@@

It may be puzzling as to why the stellar emission in the companion 
is spatially offset from RXJ1131 while its CO gas appears to be mixed-in if they 
have encountered.
This phenomenon is in fact not unique to our source. 
The gas cores of the individual 
% merging galaxies 
mergers NGC~6090 and NGC~6240 are also observed
to be coalescing  
% is in the process of or has already coalesced into a single component
while their stellar cores remain % distinct
separated \citep[0.5$-$3\,kpc;][]{Bryant99a}. 
Although this is difficult to reconcile with the fact that the stellar light in the companion 
originates from recent \SF (UV emission). 
More sensitive and higher resolution observations will be needed
to confirm any offsets between the CO gas and the stellar light in the companion.
% allow for a more detailed lens modeling to trace the gas distribution and morphology in the system
Another puzzle may be the absence of morphological features.
% tidal tails if they have already encountered.
One straightforward explanation may simply be the lack of spatial resolution.
Alternatively, it may be a result of the offset in timing between observable
morphological disturbances and the merger-induced starbursts \citep{Lotz08a}.
% with strong morphological disturbances occurring before bursts of merger-induced star formation and for shorter durations

\subsection{Fate of RXJ1131}
A classical picture of mergers is that they are responsible for the formation of local spheroids.
The growing consensus that more realistic
treatments of star formation and supernovae feedback ? which suppress
efficient early star formation ? play a critical role in enabling
disc formation in simulations.
Mergers are expected to evolve into
local ellipticals or large-disks (bulge dominated; Springel05a, Robertson06a)
and disc with low B/D ratio (Hopkins09a) from simulations.
Reason for suppression of bulge formation in mergers of gas-rich discs (even in major merger):
 because in systems that are sufficiently gas-rich, there's little/no significant (collisionless) stellar material available to torque away the cold gas in the disc; little/no angular momentum is lost, and gas will rapidly reform a cold disc in the background of the relaxing stellar potential (collisionless):
Gas, being collisional, cannot violently relax, but must
have its angular momentum torqued away in order to dissipate and
build a bulge by forming stars in a central starburst (\ie gas lose angular momnentum to stars in the perturbed disk and fall toward center). In a merger,
this torque is primarily internal, from stars in the same disc: the
passages and merger of the secondary induce a non-axisymmetric
distortion in the primary stellar and gas disc. The stellar distortion
(in a trailing resonance and close proximity to the gas response)
efficiently removes angular momentum from the gas and allows it
to fall to the center and form a bulge (see also Barnes \& Hernquist
1996; cold gas with no angular momentum will be largely unable to form any sort of disk, and that a galaxy?s worth of gas compressed to high densities and small radii will inevitably form a large mass in stars). But it no significant stellar feature is there (e.g. in the more gas-rich ones (Hopkins09a), the final product will have a bulge mass fraction $\sim$0.1$-$0.2).
i.e. small bulge disc can also form from minor merger or major merger (if gas fraction is sufficiently large)

% fate
But since the two galaxies are found to be gas-rich, can dissipates energy
through e.g. frictional forces and drive gas toward the center, leading to
higher rotational velocity and dispersion in the central region.
In that case,
it is unlikely that RXJ1131 will evolve into present-day massive ellipticals (M$\gtrsim$1\E{12}\,\Msun), since RXJ1131 is now a minor merger.
Instead, it is likely to evolve into ``normal'' ellipticals (depending on the feedback, since a major gas-rich merger without AGN feedback does not, by itself, produce a red and dead spheroidal galaxy (Lotz08a).) or massive disk galaxies where the stellar populations from the companion is dragged and added onto the dominant galaxy RXJ1131 to thicken the disc, consistent with recent models (disk from gas-rich major merger:
\citep[\eg][]{Lotz08a, Robertson06a}.

% evolve into disk
Given the
massive gas content distributed in an extended configuration/disc with its SFE on the low end of other ULIRGs, we suggest RXJ1131 may evolve into the
latter (large-disk) unless the ``companion'' somehow triggers some other
mechanisms in the latter stage, transporting the gas 
toward the central nuclear region and to simulate a final burst to form a large bulge or somehow destroy the
disc in RXJ1131.

>>>>

Alternatively, the relatively low SFE maybe intrinsic to RXJ1131.
In that case,  RXJ1131 may evolve into a disc galaxy with a small bulge component
in the context of theoretical simulations by \citet{Hopkins09a}, who suggest 
that the disc component in a gas-rich progenitor galaxy can survive post-merging
by lowering the \SF efficiency (and thus less efficient at removing the angular momentum of
the gas), thereby enabling a higher gas fraction to be retained and 
distributed over a larger extent.
%since there is little/no significant stellar material % (collisionless)
%available to torque away the cold gas in the disc, thus little/no angular momentum is lost, 
%and gas will rapidly reform a cold disc in the background 
%of the relaxing stellar potential.
% --> suppress the formation of a bulge in mergers of gas-rich discs (even in major merger)
% the final product will have a low bulge mass fraction.
%
%i.e. small bulge disc can also form from minor merger or major merger (if gas fraction is sufficiently large).
% 1:1 gas-riich mergers can yield disc-dominated remnants, and 1:3-1:4 ratio major mergers can yield B/T < 0.2.

@@

It may be puzzling as to why the stellar emission in the companion 
is spatially offset from RXJ1131 while its CO gas appears to be mixed-in if they 
have encountered.
This phenomenon is in fact not unique to our source. 
The gas cores of the individual 
% merging galaxies 
mergers NGC~6090 and NGC~6240 are also observed
to be coalescing  
% is in the process of or has already coalesced into a single component
while their stellar cores remain
separated \citep[0.5$-$3\,kpc;][]{Bryant99a}. 
Although this is difficult to reconcile with the fact that the stellar light in the companion 
originates from recent \SF (UV emission). 
More sensitive and higher resolution observations will be needed
to confirm any offsets between the CO gas and the stellar light in the companion.
% allow for a more detailed lens modeling to trace the gas distribution and morphology in the system
Another puzzle may be the absence of morphological features.
One straightforward explanation may simply be the lack of spatial resolution.
Alternatively, it may be a result of the offset in timing between observable
morphological disturbances and the merger-induced starbursts \citep{Lotz08a}.
% with strong morphological disturbances occurring before bursts of merger-induced star formation and for shorter durations

\subsection{Fate of RXJ1131}
The classical picture for mergers is one where they are responsible for the formation of the local 
red and passive spheroidal galaxies.
With more realistic treatments of star formation and feedbacks in recent simulations,
% ? which suppress efficient early star formation ? play a critical role in enabling disc formation in simulations.
it has been suggested that it is possible  to suppress bulge formation 
in gas-rich mergers, thereby evolving into large 
disks \citep{Springel05a, Robertson06a, Hopkins09a}.
% (bulge dominated; Springel05a, Robertson06a)
% and disc with low B/D ratio ($\sim$0.1$-$0.2; Hopkins09a).
%
%In systems that are sufficiently gas-rich, it is possible to 
%suppress the formation of a bulge in mergers of gas-rich discs (even in major merger)
%since there is little/no significant stellar material % (collisionless)
%available to torque away the cold gas in the disc, thus little/no angular momentum is lost, 
%and gas will rapidly reform a cold disc in the background 
%of the relaxing stellar potential.
%Gas, being collisional, cannot violently relax, but must
%have its angular momentum torqued away in order to dissipate and
%build a bulge by forming stars in a central starburst 
%(\ie gas loses angular momentum to stars in the
%perturbed disk and fall toward center).
% In a merger,
%this torque is primarily internal, from stars in the same disc: the
%passages and merger of the secondary induce a non-axisymmetric
%distortion in the primary stellar and gas disc. The stellar distortion
%(in a trailing resonance and close proximity to the gas response)
%efficiently removes angular momentum from the gas and allows it
%to fall to the center and form a bulge  (see also Barnes \& Hernquist
% 1996).
% A bulge will form because cold gas with no angular momentum 
% will be largely unable to form any sort of disk, and that a galaxy?s 
% worth of gas compressed to high densities and small
% radii will inevitably form a large mass in stars). 
%
Since both RXJ1131 and its companion are gas-rich
% unlikely RXJ1131 will evolve into present-day massive 
% ellipticals (M$\gtrsim$1\E{12}\,\Msun), 
% gas-poor mergers: --> present-day massive elliticals
% (Boylan-Kolchin, Ma & Quataert 2005, 2006; Naab, Khochfar & Burkert 2006; Robertson B., Cox T. J., Hernquist L., Franx M., Hopkins P., Martini P., Springel V., 2006a, ApJ, 641, 21; Cox et al. 2006b)
and that the bulk of the angular momentum of the gas is  
retained in minor mergers,
it is likely that RXJ1131 will evolve into either a 
``normal'' elliptical/S0 galaxy or a large disk galaxy.  
% normal E only because gE needs ways to remove angular momentum efficiently for gas to funnel
In the latter, the stellar populations from the companion will be displaced onto the 
dominant galaxy to thicken the disk \citep[\eg][]{Lotz08a, Robertson06a}.
% disk from gas-rich major merger: \citep[\eg][]{Lotz08a, Robertson06a}.
Given that the molecular gas in 
RXJ1131 is distributed in an extended disk configuration and that its 
SFE is on the low end among other ULIRGs,  % not enough stellar material to efficiently remove ang. mom.
it is more conceivable that RXJ1131 will evolve into the
latter.
% Lotz08a: a major gas-rich merger without AGN feedback does not, by itself, produce a red and dead spheroidal galaxy
Other mechanisms e.g. formation of bar-like structures will be needed to remove the
angular momentum of the gas more efficiently to transform the gas disc into a stellar spheroid and
AGN feedbacks are also necessary to halt further activities for RXJ1131 to evolve into an 
E/S0 galaxy \citep{Lotz08a}.
